---
title: "Chapter_04"
output:
  html_document:
    df_print: paged
---

```{r, message=FALSE}
library(rethinking)
```


# 4.1 Why normal distributions are normal

```{r}
pos <- replicate(1000, sum(runif(16, -1, 1)))
hist(pos)
plot(density(pos))
dens(pos)
```

```{r}
# 4.2
prod(1 + runif(12, 0, 0.1))
growth <- replicate(10000, prod(1 + runif(12, 0, 0.1)))
dens(growth, norm.comp = TRUE)

big <- replicate(10000, prod(1 + runif(12, 0, 0.5)))
dens(big, norm.comp = TRUE)

small <- replicate(10000, prod(1 + runif(12, 0, 0.01)))
dens(small, norm.comp = TRUE)
```

```{r}
# 4.5
log.big <- replicate(10000, log(prod(1 + runif(12, 0, 0.5))))
dens(log.big, norm.comp = TRUE)
```


# 4.2 A language for describing models

```{r}
# 4.6
w <- 6
n <- 9
p_grid <- seq(from = 0, to = 1, length.out = 100)
posterior <- dbinom(w, n, p_grid) * dunif(p_grid, 0, 1)
posterior <- posterior/sum(posterior)

```

# 4.3 A Gaussian model of height

```{r, message=FALSE}
library(rethinking)
data("Howell1")
d <- Howell1
d
str(d)
```

```{r}
# d$height
d2 <- d[d$age >= 18,]
dens(d2$height)
curve(dnorm(x, 178, 20), from = 100, to = 250)
curve(dunif(x, 0, 50), from = -10, to = 60)
```

```{r}
# 4.13
sample_mu <- rnorm(1e4, 178, 20)
sample_sigma <- runif(1e4, 0, 50)
prior_h <- rnorm(1e4, sample_mu, sample_sigma)
dens(prior_h)
```

## 4.3.3 Grid approximation of the posterior distribution

```{r}
## R code 4.14
mu.list <- seq( from=140, to=160 , length.out=200 )
sigma.list <- seq( from=4 , to=9 , length.out=200 )
post <- expand.grid( mu=mu.list , sigma=sigma.list )
post$LL <- sapply( 1:nrow(post) , function(i) sum( dnorm(
                d2$height ,
                mean=post$mu[i] ,
                sd=post$sigma[i] ,
                log=TRUE ) ) )
post$prod <- post$LL + dnorm( post$mu , 178 , 20 , TRUE ) +
    dunif( post$sigma , 0 , 50 , TRUE )
post$prob <- exp( post$prod - max(post$prod) )

```


```{r}
# 4.15
contour_xyz(post$mu, post$sigma, post$prob)
image_xyz(post$mu, post$sigma, post$prob)
```

## 4.3.4 Sampling from the posterior

```{r}
# 4.17
sample.rows <- sample(1:nrow(post), size = 1e4, replace = TRUE, prob = post$prob)
sample.mu <- post$mu[sample.rows]
sample.sigma <- post$sigma[sample.rows]

plot(sample.mu, sample.sigma, cex=1, pch = 16, col = col.alpha(rangi2, 0.1))

```

```{r}
dens(sample.mu)
dens(sample.sigma)
```

```{r 4.20}
HPDI(sample.mu)
HPDI(sample.sigma)
```

```{r 4.21}
d3 <- sample(d2$height, size = 20)
```

```{r 4.22}
mu.list <- seq( from=150, to=170 , length.out=200 )
sigma.list <- seq( from=4 , to=20 , length.out=200 )
post2 <- expand.grid( mu=mu.list , sigma=sigma.list )
post2$LL <- sapply( 1:nrow(post2) , function(i)
    sum( dnorm( d3 , mean=post2$mu[i] , sd=post2$sigma[i] ,
    log=TRUE ) ) )
post2$prod <- post2$LL + dnorm( post2$mu , 178 , 20 , TRUE ) +
    dunif( post2$sigma , 0 , 50 , TRUE )
post2$prob <- exp( post2$prod - max(post2$prod) )
sample2.rows <- sample( 1:nrow(post2) , size=1e4 , replace=TRUE ,
    prob=post2$prob )
sample2.mu <- post2$mu[ sample2.rows ]
sample2.sigma <- post2$sigma[ sample2.rows ]
plot( sample2.mu , sample2.sigma , cex=0.5 ,
    col=col.alpha(rangi2,0.1) ,
    xlab="mu" , ylab="sigma" , pch=16 )
```

```{r}
dens(sample2.sigma, norm.comp = TRUE)
```

## 4.3.5 Fitting the model with map
```{r 4.24}
library(rethinking)
data("Howell1")
d <- Howell1
d2 <- d[d$age >= 18,]
```

```{r 4.25}
flist <- alist(
  height ~ dnorm(mu, sigma),
  mu ~ dnorm(178, 20),
  sigma ~ dunif(0, 50)
)

m4.1 <- map(flist, data = d2)

precis(m4.1)
```

```{r 4.29}
m4.2 <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu ~ dnorm(178, 0.1),
    sigma ~ dunif(0, 50)
  ), 
  data = d2
)

precis(m4.2)
```

## 4.3.6 Sampling from a map fit
```{r 4.30}
vcov(m4.1)
```

```{r 4.31}
diag(vcov(m4.1))
cov2cor(vcov(m4.1))
```

```{r 4.32}
library(rethinking)
post <- extract.samples(m4.1, n = 1e4)
head(post)
precis(post)
```

```{r}
plot(post, cex = 0.5, pch = 16, col = col.alpha(rangi2, 0.5))
```

# 4.4 Addig a predictor

```{r}
plot(d2$height ~ d2$weight)
```

### 4.4.1 The linear model strategy
### 4.4.2 Fitting the model
```{r 4.38}
# load data again since it's a long way back
library(rethinking)
data("Howell1")
d <- Howell1
d2 <- d[d$age >= 18,]

# fit model
m4.3 <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b*weight,
    a ~ dnorm(156, 100),
    b ~ dnorm(0, 10),
    sigma ~ dunif(0, 50)
  ), 
  data = d2
)

```

### 4.4.3 Interpreting the model fit

```{r 4.40}
precis(m4.3)
```

```{r 4.41}
precis(m4.3, corr = TRUE)
```

```{r 4.42}
d2$weight.c <- d2$weight - mean(d2$weight)
```

```{r 4.43}
m4.4 <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b*weight.c,
    a ~ dnorm(178, 100),
    b ~ dnorm(0, 10),
    sigma ~ dunif(0, 50)
  ),
  data = d2
)

precis(m4.4, corr = TRUE)
```

```{r 4.45}
plot(height ~ weight, data = d2)
abline(a = coef(m4.3)["a"], b = coef(m4.3)["b"])
```


```{r 4.46}
post <- extract.samples(m4.3)
post[1:5,]
```

```{r 4.48}
N <- 10
dN <- d2[1:N, ]
mN <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b*weight,
    a ~ dnorm(178, 100),
    b ~ dnorm(0, 50),
    sigma ~ dunif(0, 50)
  ), data = dN
)
```

```{r 4.49}
# extract 20 samples from the posterio
post <- extract.samples(mN, n = 20)

# display raw data and sample size
plot(dN$weight, dN$height,
     xlim = range(d2$weight), ylim = range(d2$height),
     col = rangi2, xlab="weight", ylab="height")
mtext(concat("N = ", N))

# plot the lines with transparency
for (i in 1:20) abline(a = post$a[i], b = post$b[i], col=col.alpha("black", 0.3))

```

```{r 4.50}
mu_at_50 <- post$a + post$b * 50 
dens(mu_at_50, col = rangi2, lwd = 2, xlab = "mu|weight = 50")
```

```{r 4.52}
HPDI(mu_at_50, prob = 0.89)
```

```{r 4.53}
mu <- link(m4.3)
str(mu)
```


```{r 4.54}
# define sequence of weights to compute predictions for
# these values will be on the horizontal axis
weight.seq <- seq( from=25 , to=70 , by=1 )

# use link to compute mu
# for each sample from posterior
# and for each weight in weight.seq
mu <- link( m4.3 , data=data.frame(weight=weight.seq) )
str(mu)
```

```{r 4.55}
# use type="n" to hide raw data
plot(height ~ weight, d2, type = "n")

# loop over samples and plot each mu value
for(i in 1:100){
  points(weight.seq, mu[i,], pch = 16, col = col.alpha(rangi2, 0.1))
}
```

```{r 4.56}
mu.mean <- apply(mu, 2, mean)
mu.HPDI <- apply(mu, 2, HPDI, prob = 0.89)
```

```{r 4.57}
# plot raw data
# fading out points to make line and interval more visible
plot(height ~ weight, data = d2, col=col.alpha(rangi2, 0.5))

# plot the MAP line, aka the mean mu for each weight
lines(weight.seq, mu.mean)

# plot a shaded region for the 89% HPDI
shade(mu.HPDI, weight.seq)
```

```{r 4.59}
sim.height <- sim(m4.3, data = list(weight = weight.seq))
str(sim.height)
```

```{r 4.60}
height.PI <- apply(sim.height, 2, PI, prob=0.89)

# plot raw data
plot( height ~ weight , d2 , col=col.alpha(rangi2,0.5) )

# draw MAP line
lines( weight.seq , mu.mean )

# draw HPDI region for line
shade( mu.HPDI , weight.seq )

# draw PI region for simulated heights
shade( height.PI , weight.seq )

```

# 4.5 Polynomial regression

```{r 4.64}
library(rethinking)
data("Howell1")
d <- Howell1
str(d)

plot(d$weight, d$height)

```

```{r 4.65}
d$weight.s <- (d$weight - mean(d$weight))/sd(d$weight)
plot(d$weight.s, d$height)
```

```{r 4.66}
d$weight.s2 <- d$weight.s^2
m4.5 <- map(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b1*weight.s + b2*weight.s2 ,
        a ~ dnorm( 178 , 100 ) ,
        b1 ~ dnorm( 0 , 10 ) ,
        b2 ~ dnorm( 0 , 10 ) ,
        sigma ~ dunif( 0 , 50 )
    ) ,
    data=d )
```

```{r 4.67}
precis(m4.5)
```

```{r 4.68}
weight.seq <- seq(from = -2.2, to = 2, length.out = 30)
pred_dat <- list(weight.s = weight.seq, weight.s2 = weight.seq ^ 2)
mu <- link(m4.5, data = pred_dat)
mu.mean <- apply(mu, 2, mean)
mu.PI <- apply(mu, 2, PI, prob = 0.89)
sim.height <- sim(m4.5, data = pred_dat)
height.PI <- apply(sim.height, 2, PI, prob = 0.89)

plot(height ~ weight.s, d, col = col.alpha(rangi2, 0.5))
lines(weight.seq, mu.mean)
shade(mu.PI, weight.seq)
shade(height.PI, weight.seq)
```

```{r 4.70}
d$weight.s3 <- d$weight.s^3
m4.6 <- map(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b1*weight.s + b2*weight.s2 + b3*weight.s3 ,
        a ~ dnorm( 178 , 100 ) ,
        b1 ~ dnorm( 0 , 10 ) ,
        b2 ~ dnorm( 0 , 10 ) ,
        b3 ~ dnorm( 0 , 10 ) ,
        sigma ~ dunif( 0 , 50 )
    ) ,
    data=d )
```

```{r}
weight.seq <- seq(from = -2.2, to = 2, length.out = 30)
pred_dat <- list(weight.s = weight.seq, 
                 weight.s2 = weight.seq^2,
                 weight.s3 = weight.seq^3)
mu <- link(m4.6, data = pred_dat)
mu.mean <- apply(mu, 2, mean)
mu.PI <- apply(mu, 2, PI, prob = 0.89)
sim.height <- sim(m4.6, data = pred_dat)
height.PI <- apply(sim.height, 2, PI, prob = 0.89)

plot(height ~ weight.s, d, col = col.alpha(rangi2, 0.5))
lines(weight.seq, mu.mean)
shade(mu.PI, weight.seq)
shade(height.PI, weight.seq)
```

```{r 4.71}
plot(height ~ weight.s, d, col = col.alpha(rangi2, 0.5), xaxt = "n")
at <- c(-2, -1, 0, 1, 2)
labels <- at*sd(d$weight) + mean(d$weight)
axis(side = 1, at = at, labels = round(labels, 1))

```


# 4.7 Practice
## Easy
### 4E1
The first line

### 4E2
Two

### 4E3
p(mu, sigma | y = Pi Normal(hi|mu, sigma)Normal(mu | 0, 10) Uniform( sigma | 0, 10))/...

### 4E4
The second line

### 4E5
Three: a, b, s

## Medium
### 4M1

```{r}
sample.mu <- rnorm(1e4, 0, 10)
sample.sigma <- runif(1e4, 0, 10)
sample.y <- rnorm(1e4, sample.mu, sample.sigma)

dens(sample.y)
```

### 4M2
```{r}
flist <- alist(
  height ~ dnorm(mu, sigma),
  mu ~ dnorm(0, 10),
  sigma ~ dunif(0, 10)
)

```

### 4M3
```{r}
flist <- alist(
  y ~ dnorm(mu, sigma),
  mu <- a + b * x,
  a ~ dnorm(0, 50),
  b ~ dnorm(0, 10),
  sigma ~ dunif(0, 50)
)
```

\begin{align}
y_{i} &\sim \text{Normal}(\mu, \sigma) \\
\mu_{i} &= \alpha + \beta x_{i} \\
\alpha &\sim \text{Normal}(0, 50) \\
\beta &\sim \text{Uniform}(0, 10) \\
\sigma &\sim \text{Uniform}(0, 50)
\end{align}

___
### 4M4

\begin{align}
h_{i} &\sim \text{Normal}(\mu, \sigma) \\
\mu &= \alpha + \beta x_{i} \\
\alpha &\sim \text{Normal}(150, 50) \\
\beta &\sim \text{Normal}(4, 2) \\
\sigma &\sim \text{Uniform}(0, 50)
\end{align}

___
### 4M5

\begin{align}
h_{i} &\sim \text{Normal}(\mu, \sigma) \\
\mu &= \alpha + \beta x_{i} \\
\alpha &\sim \text{Normal}(120, 10) \\
\beta &\sim \text{Normal}(7, 1) \\
\sigma &\sim \text{Uniform}(0, 20)
\end{align}

___
### 4M6

\begin{align}
h_{i} &\sim \text{Normal}(\mu, \sigma) \\
\mu &= \alpha + \beta x_{i} \\
\alpha &\sim \text{Normal}(120, 10) \\
\beta &\sim \text{Normal}(7, 1) \\
\sigma &\sim \text{Uniform}(0, 8)
\end{align}

## Hard
### 4H1

Build the MAP model:
```{r}
library(rethinking)
data("Howell1")

d <- Howell1
m4H1 <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b * weight,
    a ~ dnorm(140, 30),
    b ~ dnorm(0, 10),
    sigma ~ dunif(0, 50)
  ),
  data = d
)

m4H1
```

Now calculate the posterior distribution of heights for each weight value in our table:
```{r}
new_weight <- c(46.95, 43.72, 64.78, 32.59, 54.63)
pred_height <- link(m4H1, data = data.frame(weight = new_weight))
str(pred_height)
```

Plot the distributions of pred_height for each weight.
```{r}
plot(height ~ weight, d, type = "n")

for (i in 1:100) {
  points(new_weight, pred_height[i,], pch = 16, col = col.alpha(rangi2, 0.5))
}
```

Calculate the mean and HPDI for the predictions:
```{r}
(pred_mean <- apply(pred_height, 2, mean))
(pred_interval <- apply(pred_height, 2, HPDI))

```

Put it into a data frame:
```{r}
df <- data.frame(
  individual = 1:5,
  weight = new_weight,
  pred_height = pred_mean,
  lower = pred_interval[1,],
  upper = pred_interval[2,]
)

df
```


### 4H2
```{r}
d <- Howell1
df <- d[d$age < 18,]
range(df$height)
summary(df$height)
dens(df$height)
HPDI(df$height)
sd(df$height)
```

#### (a)
```{r}
m4H2 <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b * weight,
    a ~ dnorm(110, 30),
    b ~ dnorm(0, 10),
    sigma ~ dunif(0, 50)
  ),
  data = df
)

precis(m4H2, corr = TRUE)
```

For every 10 units increase in weight, the model predicts a 27 unit increase in height. The model also estimates that the standard deviation for under-18s is 8.44.

```{r}
fit <- lm(height ~ weight, data = df)
summary(fit)
```

#### (b)
```{r}
# Determine the range of weights
range(df$weight)

# Define sequence of weight to predict on
weight.seq <- seq(4, 46, by = 2)

# compute mu for each sample from the posterior and for each weight we want to predict on
mu <- link(m4H2, data = data.frame(weight = weight.seq))
str(mu)
```

Summarise the distribution of mu
```{r}
mu.mean <- apply(mu, 2, mean)
mu.HPDI <- apply(mu, 2, HPDI, prob = 0.89)
```

Calculate the prediction interval
```{r}
sim.height <- sim(m4H2, data = list(weight = weight.seq))
height.PI <- apply(sim.height, 2, PI)
```


```{r}
# Plot raw data
plot(height ~ weight, df, col = col.alpha(rangi2, 0.5))

# Add map line
lines(weight.seq, mu.mean)

# Add HPDI interval for the line
shade(mu.HPDI, weight.seq)

# Add PI interval for the simulated heights
shade(height.PI, weight.seq)

```

#### (c)
I'm concerned that:

- The relationship between height and weight does not appear to be linear in the raw data

### 4H3
#### (a)
Start with model definition:

\begin{align}
h_{i} &\sim \text{Normal}(\mu_{i}, \sigma) \\
\mu_{i} &= \alpha + \beta \text{log}(w_{i}) \\
\alpha &\sim \text{Normal}(178, 100) \\
\beta &\sim \text{Normal}(0, 100) \\
\sigma &\sim \text{Uniform}(0, 50)
\end{align}

```{r}
d <- Howell1

m4H3 <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b * log(weight),
    a ~ dnorm(178, 100),
    b ~ dnorm(0, 100),
    sigma ~ dunif(0, 50)
  ),
  data = d
)

precis(m4H3, corr = TRUE)

```

Interpretation:

- the value of _a_ is -24 cm, meaning that a person with log(0) weight in kg will be -24 cm tall. This is silly, of course. For one reason, log(0) is -Inf!
- the value of _b_ is 47 cm/log(kg), meaning that for every increase of one log unit of weight, the height increases by 47 cm. 
- the estimate of $\sigma$ means that, in the model, the estimate of standard deviation of height is 5.1 cm.

#### (b)
```{r}
plot(height ~ weight, data = Howell1, col = col.alpha(rangi2, 0.4))
```

```{r}
# Estimate the MAP regression line and 89% HPDI for the mean
weight.seq <- seq(
  from = min(d$weight),
  to = max(d$weight),
  by = 2
)

mu <- link(m4H3, data = data.frame(weight = weight.seq))
str(mu)
```

Summarise the distribution of $\mu$
```{r}
mu.mean <- apply(mu, 2, mean)
mu.HPDI <- apply(mu, 2, HPDI)
```

Calculate the prediction interval
```{r}
sim.height <- sim(m4H3, data = list(weight = weight.seq))

height.PI <- apply(sim.height, 2, PI)
```

Plot it
```{r}
# Plot original data
plot(height ~ weight, data = Howell1, col = col.alpha(rangi2, 0.4))

# Add regression line for mu
lines(weight.seq, mu.mean)

# Add HPDI interval for the line
shade(mu.HPDI, weight.seq)

# Add PI interval for the predicted heights
shade(height.PI, weight.seq)
```

